---
title: "PubH7462_hw2_panxx408"
author: "Mingming Pan"
date: "2/9/2022"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(forcats)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(DataExplorer)
library(gt)
#Working directory for .RMD
knitr::opts_knit$set(echo = TRUE,
                     root.dir = rprojroot::find_rstudio_root_file())

#Controlling figure output in markdown
knitr::opts_chunk$set(
#  fig.height =   
  fig.width = 6,
#  fig.asp = .5,
  out.width = "90%",
#  out.height = 
 fig.align  = "center",
  cache = FALSE,
  eval  = TRUE,
  echo  = TRUE
)

#My Colours (from viridis)
my_purple <- "#440154FF"
my_yellow <- "#FDE725FF"

#Set Theme for ggplot2
theme_set(theme_bw() + theme(plot.title = element_text(hjust = 0.5), legend.position = "bottom"))

#Set Scientific notation output and decimal places for knitr
options(scipen = 999)
options(digits = 4)
```


```{r, message = FALSE}
#Read data
brfss_df <- read_csv("./data/brfss_smart_2010.csv")
```
## Problem3.1 Data Exploration & Cleaning
```{r}
#Data Exploration
plot_str(brfss_df)
plot_intro(brfss_df)
```

```{r, warning=FALSE}
#Date Cleaning
brfss_clean <- brfss_df %>% separate(col = Locationdesc, into = c("State", "2", "County"),
                                     sep = " ") %>%
  filter(Topic == "Overall Health") %>%
  select(Year, State, County, Response, Sample_Size, Data_value) %>%
  rename(Response_Prop = Data_value) %>%
  mutate(State = as.factor(State)) %>%
  mutate(County = as.factor(County))
```


## 3.2 Data Description
+ There are `r ncol(brfss_clean)` variables and `r nrow(brfss_clean)` observations.
+ Each observation represents details of one of the response of a county at a specific year.
+ Year, State, County, Response, Response Proportion and sample size are the six variables.
+ Year describes the year taking the questionnaire; State and County describe the locations; Response describes the type of responses and response proportion indicates the proportion of corresponding response; and sample size describes the sample size of corresponding response. 
+ The number of countys taking the questionnaire varies from year to year. 
 
 
## 3.3 Do Data Science


### 3.3.1In the year 2004, which states were observed at 6 locations?
```{r}
brfss_clean %>%
  filter(Year == "2004") %>%
  select(State, County) %>%
  distinct() %>%
  group_by(State) %>%
  summarise(Locations = n()) %>%
  filter(Locations == 6) %>%
  gt() %>%
  cols_label(Locations = "Number of Locations") %>%
  tab_header("States were Observed at 6 Locations in 2004")
```
**CO, NM, SC, TX, UT and VT were observed at 6 locations in 2004.**


### 3.3.2 Make a spaghetti plot that shows the number of observed locations in each state from 2002 to 2010. Which state has the highest mean number of locations over this period?
```{r, message=FALSE}
brfss_clean %>% filter(Year > 2001 & Year < 2011) %>%
  select(Year, State, County) %>%
  distinct() %>% 
  group_by(State, Year) %>%
  summarise(Location = n()) %>% 
  ungroup(Year, State) %>%
  mutate(State = fct_reorder(State, Location, mean, .desc = TRUE)) %>%
  ggplot() +
  geom_line(aes(x = Year, y = Location, group = State, color = State)) +
  labs(
    x = "Year",
    y = "Number of Locations",
    title = "Mean of Locations by State from 2002 - 2010"
  )
```

**NJ is the state which has the highest mean number of locations over the period.**


### 3.3.3 Make a table showing, for the years 2002, 2006, and 2010, the mean and standard deviation of sample size and proportion of Excellent, Good, and Poor responses across locations in MN.
```{r, message=FALSE}
brfss_clean %>% filter(Year %in% c("2002", "2006", "2010"),
                       State == "MN",
                       Response %in% c("Excellent", "Good", "Poor")) %>%
  group_by(Response, Year) %>%
  summarise(across(c(Sample_Size, Response_Prop), 
                   list(mean = mean,
                        sd = sd),
                   na.rm = FALSE,
                   .names = "{.col}.{.fn}")) %>%
    gt() %>%
  tab_header("Response Proportion and Sample Size in MN on 2002, 2006 and 2010") 
```

**The response proportions in there years do not have obvious variance, while the sample size is increase from 2002 to 2010.**


### 3.3.4 Create a ggplot that communicates the results/trends from the table above and stands on its own.
```{r, message=FALSE, warning=FALSE}
brfss_clean %>% filter(Year %in% c("2002", "2006", "2010"),
                       State == "MN",
                       Response %in% c("Excellent", "Good", "Poor")) %>%
  group_by(Response, Year) %>%
  summarise(across(c(Sample_Size, Response_Prop), 
                   list(mean = mean,
                        sd = sd),
                   na.rm = FALSE,
                   .names = "{.col}.{.fn}")) %>%
  pivot_longer(cols = c(Sample_Size.mean, Sample_Size.sd, Response_Prop.mean, Response_Prop.sd),
               names_to = "Stat",
               values_to = "Value") %>%
  ggplot(aes(x = Year, y = Value, group = Response, color = Response)) +
  geom_point() +
  geom_line() +
  facet_wrap(~Stat) +
  labs(x = "Value",
       y = "Statistics",
       title = "Response Proportion and Sample Size in MN on 2002, 2006 and 2010")
```





